\chapter{Network Science and Complex Networks}
\label{ch:networkscience}

Network science extends classical graph theory to the study of large, real-world networks.  
Such networks often exhibit properties like clustering, heavy-tailed degree distributions, and small-world behavior \cite{newman2010networks}.

% ---------------------------------------------------------
\section{Random Graph Models}

\subsection{Erdős–Rényi Model}

In the $G(n,p)$ model, each pair of vertices is connected with probability $p$.  
This simple model facilitates theoretical analysis but rarely matches real data.

\subsection{Watts–Strogatz Model}

Designed to reproduce the ``small-world'' phenomenon:  
high clustering and short path lengths are typical of social systems.

\subsection{Barabási–Albert Model}

Models scale-free networks where degree distributions follow a power law:
\[
P(k) \propto k^{-3}.
\]

Such networks arise in the web graph, biological systems, and city transportation \cite{barabasi2016network}.

% ---------------------------------------------------------
\section{Centrality Measures}

Centrality quantifies importance of nodes in a network.

\subsection{Degree Centrality}

Measures local influence:
\[
C_D(v) = \deg(v).
\]

\subsection{Closeness Centrality}

Measures how close a node is to others:
\[
C_C(v) = \frac{1}{\sum_{u} d(v,u)}.
\]

\subsection{Betweenness Centrality}

Measures how often a node lies on shortest paths:
\[
C_B(v) = \sum_{s,t} \frac{\sigma_{st}(v)}{\sigma_{st}}.
\]

\subsection{Eigenvector Centrality and PageRank}

Eigenvector centrality measures influence recursively:  
high-scoring nodes connect to high-scoring neighbors.  
PageRank modifies this with damping for web ranking \cite{brin1998anatomy}.

% ---------------------------------------------------------
\section{Community Detection}

Communities are densely connected node groups.

\subsection{Modularity}

Measures how well a partition captures community structure:
\[
Q = \frac{1}{2m} \sum_{ij} (A_{ij} - \frac{k_i k_j}{2m})\delta(c_i,c_j).
\]

Algorithms such as Louvain and Leiden optimize modularity.

% ---------------------------------------------------------
\section{Summary}

This chapter introduced models and measures that help analyze real network behavior.  
Network science provides tools essential for identifying influential nodes, community structure, and overall graph organization.