\chapter*{Abstract}
\addcontentsline{toc}{chapter}{Abstract}

This report provides a comprehensive exploration of graph theory and its pivotal role in modern data science and artificial intelligence. Beginning with fundamental mathematical concepts, we systematically progress through classical graph algorithms to advanced machine learning techniques, culminating in practical applications across various domains.

The first section establishes the mathematical foundations, covering essential definitions, graph types, and representations. We then examine classical algorithms including traversal methods, shortest path algorithms, and network flow techniques. The core of this report focuses on the intersection of graph theory and machine learning, with detailed coverage of Graph Neural Networks (GNNs), their architectures, and training methodologies.

Practical applications are explored through case studies in social network analysis, recommendation systems, natural language processing, computer vision, fraud detection, and biological sciences. The report also addresses scalability considerations through discussions of graph databases and distributed processing frameworks.

Key contributions include a comparative analysis of graph algorithms, implementation guidelines for GNNs, and evaluation of graph-based approaches against traditional methods. The findings demonstrate that graph theory provides not only theoretical foundations but also practical tools for extracting meaningful insights from interconnected data structures.

\vspace{0.5cm}
\noindent\textbf{Keywords:} Graph Theory, Data Science, Artificial Intelligence, Graph Neural Networks, Network Analysis, Machine Learning, Algorithms, Mathematical Foundations